---
title: "Lab10"
author: "Tristan LaRose"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

## Get WD

```{r}
getwd()
```

# Task 2

## Mymaxlik()

### A

A matrix is created with the outer function. It applies the lfun function to x and param, which is all then put into z. Each row is the product of an x-value and all the params. Each column is a product of a value of param and all the x values.

### B

This line finds the index of the y that is the biggest. If there are more than one, it will take the last of them.

### Maximum Likelihood

#### Formula

$P(y_1y_2y_3y_4y_5y_6y_7y_8)=p(y_1)×p(y_2)×p(y_3)×p(y_4)×p(y_5)×p(y_6)×p(y_7)×p(y_8)=∏^8_i=1(^{20}_{y_i})y_ip^{y_i}(1−p)^{20−y_i}$

#### R-code

$∏^8_{i=1}dbinom(y_i,size=20,prob=p)=dbinom(y_1,size=20,prob=p)×dbinom(y_2,size=20,prob=p)×dbinom(y_3,size=20,prob=p)×dbinom(y_4,size=20,prob=p)×dbinom(y_5,size=20,prob=p)×dbinom(y_6,size=20,prob=p)×dbinom(y_7,size=20,prob=p)×dbinom(y_8,size=20,prob=p)$

#### Mamimum Likelihood for $y=3,3,4,3,4,5,5,4$

```{r}
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix – try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun) # A
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y >= max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y))) # B
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope should change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}
logbin=function(x,param) log(dbinom(x,prob=param,size=20))
mymaxlik(x=c(3,3,4,3,4,5,5,4),param=seq(0,1,length=1000),lfun=logbin,xlab=expression(pi),main="Binomial",cex.main=2)
```

The maximum likelihood is 0.19

# Task 3

## $y=4,6,7,6,5$

```{r}
logpoiss=function(x,param) log(dpois(x,lambda=param))
mymaxlik(x=c(4,6,7,6,5),param=seq(0,10,length=1000),lfun=logpoiss,xlab=expression(lambda),main="Poisson",cex.main=2)
```

### Algebraic Expression

$l(λ)=−5λ+∏^5_{i=1}Y_ilogλ−∏^5_{i=1}log(Y_i!)$

### Maximum Likelihood

$\hat\lambda =5.600000$

## MyNRML

```{r}
myNRML=function(x0,delta=0.001,llik,xrange,parameter="param"){
f=function(x) (llik(x+delta)-llik(x))/delta
fdash=function(x) (f(x+delta)-f(x))/delta
d=1000
i=0
x=c()
y=c()
x[1]=x0
y[1]=f(x[1])
while(d > delta & i<100){
i=i+1
x[i+1]=x[i]-f(x[i])/fdash(x[i])
y[i+1]=f(x[i+1])
d=abs(x[i+1]-x[i])
}
layout(matrix(1:2,nr=1,nc=2,byrow=TRUE),width=c(1,2))
curve(llik(x), xlim=xrange,xlab=parameter,ylab="log Lik",main="Log Lik")
curve(f(x),xlim=xrange,xaxt="n", xlab=parameter,ylab="derivative",main=  "Newton-Raphson Algorithm \n on the derivative")
points(x,y,col="Red",pch=19,cex=1.5)
axis(1,x,round(x,2),las=2)
abline(h=0,col="Red")

segments(x[1:(i-1)],y[1:(i-1)],x[2:i],rep(0,i-1),col="Blue",lwd=2)
segments(x[2:i],rep(0,i-1),x[2:i],y[2:i],lwd=0.5,col="Green")

list(x=x,y=y)
}
myNRML(x0=1,delta=0.000001,llik=function(x) log(dpois(4,x)*dpois(6,x)*dpois(7,x)*dpois(6,x)*dpois(5,x)),xrange=c(0,30),parameter="lambda" )
```

# Task 4

## Biased Coin

```{r}
logbin2=function(theta){log(dbinom(2,prob=theta,size=6)) + log(dbinom(4,prob=theta,size=10))}

mymaxlikg=function(lfun="logbin2",theta) { 
nth=length(theta)  
thmat=matrix(theta,nr=nth,nc=1,byrow=TRUE) 
z=apply(thmat,1,lfun) 
zmax=max(which(z==max(z)))  
plot(theta,exp(z),type="l") 
abline(v=theta[zmax],col="Blue")   
axis(3,theta[zmax],round(theta[zmax],4))   
theta[zmax]   
}
mymaxlikg(theta=seq(0,1,length=10000),lfun=logbin2)
```

### Max Likelihood is 0.375

# Task 5

## $p(y_1,y_2│θ_1 θ_2 )=bin(y_1│θ_1 )poiss(y_2 |θ_2)$

### Algebraic Expression

$l(θ_1,θ_2)=log((ny_1)+y_1log(θ_1)+(n−y_1)log(1−θ_1)−2θ_2+y_2log(θ_2)−log(y_2!)$

```{r}
logbinpois=function(theta1,theta2) log(dbinom(4,size=20,prob=theta1)) + log(dpois(4,lambda=theta2))

maxlikg2=function(theta1,theta2,lfun="logbinpois",...){
n1=length(theta1)
n2=length(theta2)
z=outer(theta1,theta2,lfun)
contour(theta1,theta2,exp(z),...) 
maxl=max(exp(z))    
coord=which(exp(z)==maxl,arr.ind=TRUE) 
th1est=theta1[coord[1]] 
th2est=theta2[coord[2]]
abline(v=th1est,h=th2est)
axis(3,th1est,round(th1est,2))
axis(4,th2est,round(th2est,2),las=1)
list(th1est=th1est,th2est=th2est)
}
maxlikg2(theta1=seq(0,1,length=1000),theta2=seq(0,10,length=1000),nlevels=20)
```

# Task 6

##  $\hat\mu$ and $\hat\sigma$ when $y=10,12,13,15,12,11,10$ and $Y_i$~$N(\mu,\sigma)$

```{r}
mymlnorm=function(x,mu,sig,...){  
nmu=length(mu) 
nsig=length(sig)
n=length(x) 
zz=c()  
lfun=function(x,m,p) log(dnorm(x,mean=m,sd=p))   
for(j in 1:nsig){
z=outer(x,mu,lfun,p=sig[j]) 
y=apply(z,2,sum)
zz=cbind(zz,y)
}
maxl=max(exp(zz))
coord=which(exp(zz)==maxl,arr.ind=TRUE)
maxlsig=apply(zz,1,max)
contour(mu,sig,exp(zz),las=3,xlab=expression(mu),ylab=expression(sigma),axes=TRUE,
main=expression(paste("L(",mu,",",sigma,")",sep="")),...)
mlx=round(mean(x),2)  
mly=round(sqrt((n-1)/n)*sd(x),2)
abline(v=mean(x),lwd=2,col="Green")
abline(h=sqrt((n-1)/n)*sd(x),lwd=2,col="Red")

muest=mu[coord[1]]
sigest=sig[coord[2]]

abline(v=muest, h=sigest)
return(list(x=x,coord=coord,maxl=maxl))
}

mymlnorm(x=c(10,12,13,15,12,11,10),mu=seq(10,15,length=1000),sig=seq(0.1,4,length=1000),lwd=2,labcex=1)
```

# Task 7

## $sam=rbeta(30,shape1=3,shape2=4)$

```{r}
mymlbeta=function(x,alpha,beta,...){ 
na=length(alpha) 
nb=length(beta)
n=length(x) 
zz=c()    
lfun=function(x,a,b) log(dbeta(x,shape1=a,shape2=b))   
for(j in 1:nb){
z=outer(x,alpha,lfun,b=beta[j])
y=apply(z,2,sum)
zz=cbind(zz,y)
}
maxl=max(exp(zz))    
coord=which(exp(zz)==maxl,arr.ind=TRUE) 
aest=alpha[coord[1]] 
best=beta[coord[2]]
contour(alpha,beta,exp(zz),las=3,xlab=expression(alpha),ylab=expression(beta),axes=TRUE,
main=expression(paste("L(",alpha,",",beta,")",sep="")),...)

abline(v=aest, h=best)
points(aest,best,pch=19)
axis(4,best,round(best,2),col="Red")
axis(3,aest,round(aest,2),col="Red")
return(list(x=x,coord=coord,maxl=maxl,maxalpha=aest,maxbeta=best))
}
library(s20x)
layout(matrix(1:12,nr=3,nc=4,byrow=TRUE))
z=c()
a=3
b=4
sam= rbeta(30,shape1=a,shape2=b)
nsam=length(sam)
for(i in 1:12){
w=mymlbeta(x=sample(sam,nsam,replace=TRUE),alpha=seq(0.1,20,length=100),beta=seq(0.1,20,length=100),lwd=2,labcex=1,col="steelblue")
points(a,b,col="Red",pch=19)
abline(v=a,h=b,col="Red")
z<-rbind(z,c(w$maxalpha,w$maxbeta))
 }
```

```{r}
colnames(z)=c("alpha","beta")
z
```

```{r}
dz=apply(z-c(a,b),1,function(x) sqrt(sum(x^2)) )
plot(dz)
```

# Task 8

```{r}
MATH4753TRISTAN2023::myNRML(x0=2,delta=0.0001,llik=function(x) log(dpois(4,x)*dpois(6,x)*dpois(7,x)*dpois(6,x)*dpois(5,x)),xrange=c(0,30),parameter="lambda" )


```









































