---
title: "Lab9"
author: "Tristan LaRose"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

## Get WD

```{r}
getwd()
```

# Task 2

## Line A

Line A is taking sampling with replacement n*iter times. This means that every time something is sampled from the data, it is then replaced before the next sample is taken. 

## Line B

Line B is forming a confidence interval, so we can say with "confidence" that the mean of the values in the xstat vector is between $\alpha/2$ and $1-\alpha/2$.

## Sample

It is important that the sample has equally likely members because the sample is assumed to be representative of the whole population. If there is a sampling bias, it cannot be assumed that the sample is represented.

```{r}
set.seed(35) # This will give everyone the same sample
sam=round(rnorm(20,mean=10,sd=4),2)

unique(sample(sam,20,replace=TRUE) ) # repeat this line 5Xs
unique(sample(sam,20,replace=TRUE) ) # repeat this line 5Xs
unique(sample(sam,20,replace=TRUE) ) # repeat this line 5Xs
unique(sample(sam,20,replace=TRUE) ) # repeat this line 5Xs
unique(sample(sam,20,replace=TRUE) ) # repeat this line 5Xs
```

I see that they will all have the same sample process from the repeated code, but they will return unique samples each time. They aren't even all the same length.

```{r}
#sample(sam,21,replace=FALSE)
```

When I run this code, I get the following message: 
Error in sample.int(length(x), size, replace, prob) :
cannot take a sample larger than the population when 'replace = FALSE'

This is because there were only 20 taken from the rnorm function called earler, so when we call for 21, it is larger than the population.


# Task 3

## myboot

```{r}
myboot2<-function(iter=10000,x,fun="mean",alpha=0.05,cx=1.5,...){  #Notice where the ... is repeated in the code
n=length(x)   #sample size

y=sample(x,n*iter,replace=TRUE)
rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
# A histogram follows
# The object para will contain the parameters used to make the histogram
para=hist(xstat,freq=FALSE,las=1,
main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),
...)

#mat will be a matrix that contains the data, this is done so that I can use apply()
mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)

#pte is the point estimate
#This uses whatever fun is
pte=apply(mat,2,fun)
abline(v=pte,lwd=3,col="Black")# Vertical line
segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
text(ci[1],0,paste("(",round(ci[1],2),sep=""),col="Red",cex=cx)
text(ci[2],0,paste(round(ci[2],2),")",sep=""),col="Red",cex=cx)

# plot the point estimate 1/2 way up the density
text(pte,max(para$density)/2,round(pte,2),cex=cx)

invisible(list(ci=ci,fun=fun,x=x))# Some output to use if necessary
}
```

## A, B, C, and D $\alpha=0.05, iter=10000$
### A

```{r}
set.seed(39)
sam=rnorm(25,mean=25,sd=10)

myboot2(x=sam,fun="mean")
```



### B

```{r}
set.seed(30)
sam=rchisq(20,df=3)

myboot2(x=sam,fun="mean")
```

### C

```{r}
set.seed(40)
sam=rgamma(30,shape=2,scale=3)

myboot2(x=sam,fun="mean")
```

### D

```{r}
set.seed(10)
sam=rbeta(20,shape1=3,shape2=4)

myboot2(x=sam,fun="mean")
```


## A, B, C, and D $\alpha=0.20, iter=10000$

### A

```{r}
set.seed(39)
sam=rnorm(25,mean=25,sd=10)

myboot2(alpha=0.20,x=sam,fun="sd")
```

### B

```{r}
set.seed(30)
sam=rchisq(20,df=3)

myboot2(alpha=0.20,x=sam,fun="sd")
```

### C

```{r}
set.seed(40)
sam=rgamma(30,shape=2,scale=3)

myboot2(alpha=0.20,x=sam,fun="sd")
```

### D

```{r}
set.seed(10)
sam=rbeta(20,shape1=3,shape2=4)

myboot2(alpha=0.20,x=sam,fun="sd")
```

# Task 4

## Modified myboot

```{r}
myboot3<-function(iter=10000,x,fun="mean",alpha=0.05,cx=1.5,...){  #Notice where the ... is repeated in the code
n=length(x)   #sample size

y=sample(x,n*iter,replace=TRUE)
rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
# A histogram follows
# The object para will contain the parameters used to make the histogram
para=hist(xstat,freq=FALSE,las=1,
main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),
...)

#mat will be a matrix that contains the data, this is done so that I can use apply()
mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)

#pte is the point estimate
#This uses whatever fun is
pte=apply(mat,2,fun)
abline(v=pte,lwd=3,col="Black")# Vertical line
segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
text(ci[1],0,paste("(",round(ci[1],2),sep=""),col="Red",cex=cx)
text(ci[2],0,paste(round(ci[2],2),")",sep=""),col="Red",cex=cx)

# plot the point estimate 1/2 way up the density
text(pte,max(para$density)/2,round(pte,2),cex=cx)

invisible(list(ci=ci,fun=fun,x=x))# Some output to use if necessary

returnxstat <- xstat[1:30]

barplot(returnxstat)

returnxstat
}
```

```{r}
sam=c(1,1,1,2,2,2,2,3,3,3,4,4) 
myboot3(x=sam,fun="median")
```

(L,U) for the median is (1.5,3)

# Task 5

## mean/median

```{r}
mdm <-function(x,mu,med) {
  
  
  mu = mean(x)
  med = median(x)
  
  mu/med
}
set.seed(39)
sam=rnorm(25,mean=25,sd=10)

mdm(x=sam,mu=25,med=23)
```

#### $\alpha=0.05$

### A

```{r}
set.seed(39)
sam=rnorm(25,mean=25,sd=10)



myboot3(x=sam,fun="mdm")
```

### B

```{r}
set.seed(30)
sam=rchisq(20,df=3)

myboot3(x=sam,fun="mdm")
```

### C

```{r}
set.seed(40)
sam=rgamma(30,shape=2,scale=3)

myboot3(x=sam,fun="mdm")
```

### D

```{r}
set.seed(10)
sam=rbeta(20,shape1=3,shape2=4)

myboot3(x=sam,fun="mdm")
```

#### $\alpha=0.3$

### A

```{r}
set.seed(39)
sam=rnorm(25,mean=25,sd=10)



myboot3(alpha=0.3,x=sam,fun="mdm")
```

### B

```{r}
set.seed(30)
sam=rchisq(20,df=3)

myboot3(alpha=0.3,x=sam,fun="mdm")
```

### C

```{r}
set.seed(40)
sam=rgamma(30,shape=2,scale=3)

myboot3(alpha=0.3,x=sam,fun="mdm")
```

### D


```{r}
set.seed(10)
sam=rbeta(20,shape1=3,shape2=4)

myboot3(alpha=0.3,x=sam,fun="mdm")
```

# Task 6

## new distributions

### for mean

```{r}
sam=rpois(20,lambda = 2)

myboot3(alpha=0.2,x=sam,fun="mean")
```

```{r}
sam=rt(20,1,1)

myboot3(alpha=0.2,x=sam,fun="mean")
```

```{r}
sam=runif(20,min=0,max=1)

myboot3(alpha=0.2,x=sam,fun="mean")
```

```{r}
sam=rweibull(20,shape=2,scale=1)

myboot3(alpha=0.2,x=sam,fun="mean")
```

### for variance

```{r}
sam=rpois(20,lambda = 2)

myboot3(alpha=0.2,x=sam,fun="var")
```

```{r}
sam=rt(20,1,1)

myboot3(alpha=0.2,x=sam,fun="var")
```

```{r}
sam=runif(20,min=0,max=1)

myboot3(alpha=0.2,x=sam,fun="var")
```

```{r}
sam=rweibull(20,shape=2,scale=1)

myboot3(alpha=0.2,x=sam,fun="var")
```

# Task 7

## 2 interesting statisctics

```{r}
set.seed(68)
sam=rnorm(20,mean=10,sd=4) 

myboot3(x=sam,fun="IQR")
```


```{r}
set.seed(68)
sam=rnorm(20,mean=10,sd=4) 

myboot3(x=sam,fun="quantile")
```


## checking with myboot

```{r}
set.seed(68)
sam=rnorm(20,mean=10,sd=4)

myboot3(x=sam)
```

# Task 8

```{r}
ddt<-read.csv("DDT.csv")
MATH4753TRISTAN2023::myboot(x=ddt$DDT)
```


























