---
title: "Lab 1"
author: "Tristan LaRose"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks 

Lab 1 tasks

## Task 1

working directory

```{r}
getwd()
```

## Task 2

Read in the DDT data


```{r}
ddt <- read.csv(file = "DDT.csv")
head(ddt, 6)
```

## Task 3

Quantitative and Qualitative Variables

```{r}
names(ddt)
class(ddt)
summary(ddt)
```

The qualitative variables are "RIVER," "SPECIES," and "MILE."

The quantitative variables are "LENGTH," "WEIGHT," and "DDT."



How many Species?

```{r}
table(ddt$SPECIES)
```

There are three species in the ddt data set.



Largemouth Bass with weight greater than 800gms.

```{r}
ddt[ddt$SPECIES == "LMBASS" & ddt$WEIGHT > 800, ]
```



SCM River with DDT greater than 4.0.

```{r}
ddt[ddt$RIVER == "SCM" & ddt$DDT > 4.0, ]
```

## Clicker Questions

Find the mean length of fish in the ddt data frame:

```{r}
mean(ddt$LENGTH)
```



What is the standard deviation of the weight of fish in the ddt data frame?

```{r}
sd(ddt$LENGTH)
```



Is this the plot of LENGTH vs WEIGHT from the ddt data frame?

```{r}
plot(ddt$WEIGHT, ddt$LENGTH)
```

0=No. The graph from the clicker question has its axes flipped, though, it is labelled correctly.



If v=1:20 what is the last value of v/20?

```{r}
v = 1:20
v
v/20
```

1.00

## Task 4

Rivers

```{r}
riv <- table(ddt$RIVER)
riv
```

```{r}
barplot(riv, axes = TRUE, ylab = "NUMBER OF ENTRIES", xlab = "RIVER", col = c("red", "green", "blue", "orange"), main = "Number of Entries at Each River")
```

Rivers vs. Species of Fish

```{r}
rivf <- table(ddt$RIVER, ddt$SPECIES)
rivf
```

```{r}
barplot(rivf, axes = TRUE, 
        ylab = "NUMBER OF FISH", 
        xlab = "SPECIES",
        col = c("red", "green", "blue", "orange"),
        main = "River vs. Species"
        )
legend(1.5, 
       90, 
       legend = c("FCM", "LCM", "SCM", "TRM"), 
        col = c("red", "green", "blue", "orange"), 
        lty = 1:1,
        cex = 0.8
        )
```


## Task 5

Pie chart of SPECIES

```{r}
sp <- with(ddt,table(SPECIES))
sp
pie(sp, col = 2:4, main = "SPECIES")
```


Pie Chart of RIVER

```{r}
ri <- with(ddt, table(RIVER))
ri
pie(ri, col = 2:5, main = "Fish by RIVER")
```


## Task 6

A box plot for DDT

```{r}
boxplot(ddt$DDT, col = "red", ylab = "DDT", main = "DDT")
```

A boxplot for WEIGHT

```{r}
boxplot(ddt$WEIGHT, col = "red", ylab = "WEIGHT", main = "WEIGHT")
```

A boxplot for LENGTH

```{r}
boxplot(ddt$LENGTH, col = "red", ylab = "LENGTH", main = "LENGTH")
```


## Task 7

Coplot of LENGTH Vs WEIGHT given RIVER

```{r}
fishcol <- with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot(LENGTH~WEIGHT|RIVER,data=ddt,col=fishcol)
```


Coplot of DDT Vs WEIGHT given SPECIES

```{r}
fishcol <- with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot(DDT~WEIGHT|SPECIES,data=ddt,col=fishcol)
```


## Task 8

```{r}
library(ggplot2)
g <- ggplot(ddt, aes(x = SPECIES, y = WEIGHT, fill = RIVER))
g <- g + geom_boxplot() + ggtitle("Tristan LaRose")
g
```


```{r}
library(ggplot2)
g <- ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES))
g <- g + geom_violin() + ggtitle("Tristan LaRose")
g

```





```{r}
library(ggplot2)

g <- ggplot(ddt, aes(x = WEIGHT, y = LENGTH, col = SPECIES))
g <- g + geom_point() + ggtitle("Tristan LaRose")
g

```



