---
title: "Lab 5"
author: "Tristan LaRose"
date: "`r Sys.Date()`"
output: 
  html_document: 
      toc: yes
      toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Task 1

## Working Directory

```{r}
getwd()
```

# Task 2

##mybin

```{r}
mybin = function(iter = 1000, n = 10, p = 0.5) {
  sam.mat = matrix(NA, nrow = n, ncol = iter, byrow = TRUE)
  
  succ = c()
  
  for( i in 1:iter) {
    sam.mat[,i] = sample(c(1, 0), n, replace = TRUE, prob = c(p, 1 - p))
    
    succ[i] = sum(sam.mat[,i])
  }
  
  succ.tab = table(factor(succ, levels = 0:n))
  
  iter.lab = paste0("iter = ", iter)
  n.lab = paste0("n = ", n)
  p.lab = paste0("p = ", p)
  lab = paste(iter.lab, n.lab, p.lab, sep = ", ")
  barplot(succ.tab / (iter), col = rainbow(n + 1), main = "Binomial Simulation", sub = lab, xlab = "Number of Successes")
  succ.tab / iter
}
```

## iter = 100, n = 10, p = 0.7

```{r}
bin100 <- mybin(iter = 100, n = 10, p = 0.7)
```

## iter = 200, n = 10, p = 0.7

```{r}
bin200 <- mybin(iter = 200, n = 10, p = 0.7)
```

## iter = 500, n = 10, p = 0.7

```{r}
bin500 <- mybin(iter = 500, n = 10, p = 0.7)
```

## iter = 1000, n = 10, p = 0.7

```{r}
bin1000 <- mybin(iter = 1000, n = 10, p = 0.7)
```

## iter = 10000, n = 10, p = 0.7

```{r}
bin10000 <- mybin(iter = 10000, n = 10, p = 0.7)
```

## iter = 10000 Table

```{r}
bin10000
```

## Verification

```{r}
dbin.tab <- round(dbinom(0:10, size = 10, prob = 0.7), 4)
names(dbin.tab) = 0:10
dbin.tab
```

# Task 3

## Sample Without Replacement

```{r}
sample(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0), size = 5, replace = FALSE, prob = NULL)
```

## Sample With Replacement

```{r}
sample(c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0), size = 5, replace = TRUE, prob = NULL)
```

```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
  
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  
  succ=c()
  
  for( i in 1:iter){
    
    sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
    
    succ[i]=sum(sam.mat[,i])
  }
  
  succ.tab=table(factor(succ,levels=0:n))
  
  barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
  
succ.tab/iter
}
```

## iter = 100, N = 20, r = 12, n = 5

```{r}
mh1 = myhyper(iter = 100, N = 20, r = 12, n = 5)
```

## iter = 200, N = 20, r = 12, n = 5

```{r}
mh2 = myhyper(iter = 200, N = 20, r = 12, n = 5)
```

## iter = 500, N = 20, r = 12, n = 5

```{r}
mh5 = myhyper(iter = 500, N = 20, r = 12, n = 5)
```

## iter = 1000, N = 20, r = 12, n = 5

```{r}
mh10 = myhyper(iter = 1000, N = 20, r = 12, n = 5)
```

## iter = 10000, N = 20, r = 12, n = 5

```{r}
mh100 = myhyper(iter = 10000, N = 20, r = 12, n = 5)
```

## iter = 10000, N = 20, r = 12, n = 5 table

```{r}
mh100
```

## Verification

```{r}
dhyp.tab = round(dhyper(x = 0:5, m = 12, n = 8, k = 5), 4)
names(dhyp.tab) = 0:5
dhyp.tab
```

The values after 10000 iterations and from dhyper are approximately the same.

# Task 4

## What mysample should do

It appears to make a random sample n out of 1 to 10 by an amount of times described in iter. Each time it is ran it also produces a barplot. The time aspect is the time between executions. 

## Running mysample

```{r}
mysample=function(n, iter=10,time=0.5){
for( i in 1:iter){

s=sample(1:10,n,replace=TRUE)

sf=factor(s,levels=1:10)

barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)


Sys.sleep(time)
}
}

mysample(n=1000, iter=30,time=1)
```

Running it shows a new barplot every second. My computer is not strang enough to show them in real-time so I had to look through each plot after. These plots show the distributions for each sample, and they are all valued at about 0.1.

# Task 5

## 8 choose 4 

```{r}
choose(8,4)
```

## 	P(Y>4),Y~Pois(λ=2)

```{r}
1 - ppois(q = 3, lambda = 2)
```

## P(Y=10),Y~NegBin(p=0.4,r=3) 

```{r}
choose(10 - 1, 3 - 1) * 0.4 ^ 3 * 0.6 ^ (10 - 3)
```

```{r}
nbp = dnbinom(x = 10 - 3, size = 3, prob = 0.4)
nbp
```

## 	P(Y≤8),Y~Bin(n=15,p=0.4)

```{r}
pbinom(q = 8, size = 15, prob = 0.4)
```


# Task 6

```{r}
MATH4753TRISTAN2023::mhyper()
```
