---
title: "ASS3"
author: "Tristan LaRose"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 16/16

## 1

```{r}
round((exp(-120/95)),5)
```

```{r}
PHI<-read.csv("PHISHING.csv")
summary(PHI)
```

The data does seem to follow an exponential of mean 95. The sample sample mean is equal to 95.524 which is very close to 95.

## 2

```{r}
3*(0.07)
```
Mean = 0.21

```{r}
3*(0.07^2)
```
variance = 00.147

```{r}
sqrt(0.0147)

.21+3*(0.1212436)
```

Since 0.60 is outside 3 standard deviations of the mean, the gamma distribution with values a=3 and $\beta$=0.07 does not model the maximum flood value of the river.


## 3

### a

```{r}
2*2
```
Mean length of A = 4

```{r}
1*4
```
Mean length of B = 4

### b

```{r}
2*(2^2)
```
Variance of A = 8

```{r}
1*(4^2)
```
Variance of B = 16

### c

#### For A

```{r}
-0.3033-exp(-.5)+exp(0)
```

```{r}
-exp(-.5)
```

```{r}
-0.3033-0.6065307+1
```

#### For B

```{r}
-exp(-1/4)
```

```{r}
exp(0)
```

```{r}
1-0.7788008
```

Formula B has a higher probability of causing a human reaction in less then 1 minute because 0.2211992>0.0901693.

## 4

### a

$F(2)=1-e(-2^2/4)$

```{r}
1-exp(-2^2/4)
```

### b

Mean = 1.772454

```{r}
4^(1)*(gamma(2)-gamma(1.5))
gamma(2)
gamma(1.5)*gamma(1.5)

1-0.7853982
4*(0.2146018)
sqrt(0.8584072)
```
Standard deviation = 0.9265027



### c

```{r}
1.772454-2*(0.9265027)
1.772454+2*(0.9265027)

```
(-0.0805514,3.625459)



### d

```{r}
1-(1-exp(-6^2/4))
```


P(Y>6) = 0.0001234098. This is very small probability, so it is not very likely that Y will exceed 6 years.


## 5

### a

E(Y)= a/(a+B)= Mean = 0.1818182

```{r}
(2)/(2+9)
```

$\sigma^2= (aB)/(a+B)^2(a+B+1)= Variance = 0.01239669$

```{r}
(2*9)/((2+9)^2*(2+9+1))
```



### b

```{r}
1-pbeta(.40,2,9)
```

```{r}
round(1-pbeta(.40,2,9),4)
```


The probability that any randomly selected year that 40 percent is used to repair the laser color printer = 0.0464.

### c

```{r}
round(pbeta(.1,2,9),4)
```

The probability that at most 10 percent of the budget is used to repair the laser color printer = 0.2639.


## 6

### a

$\alpha=2$ and $\beta = 16$


### b

```{r}
16^(1/2)

gamma(3/2)

4*(0.8862269)
```

Mean = 3.544908

```{r}
16^(2/2)

gamma(2)-(gamma(3/2))^2

16*(0.2146018)
```

Variance = 3.433629


### c

```{r}
1-pweibull(6,2,4)

round(1-pweibull(6,2,4),4)
```

The probability a new memory chip will not fail before 6 years = 0.1054.


## 7

### a

$p(x,y)= 1/36$ $1=<x=<6,$ $1=<y=<6$


### b

$P(X=1)=p_1(1)=1/6$

The same is true for $p_1(2)= p_1(3)…=p_1(6) = 1/6$

The same is also true for $p_2(1)=p_2(2)…=p_2(6)=1/6$

```{r}
1/36+1/36+1/36+1/36+1/36+1/36
1/6
```

### c

$p_1(1|1)= p(1,1)/p_2(1)=1/6$

which is true for for every $p_1(x,y),$ $1=<x=<6,$ $1=<y=<6$
and $p_2(x,y),$ $1=<x=<6,$ $1=<y=<6$

```{r}
(1/36)/(1/6)
1/6
```

### d

$p_1(x)=p_1(x|y)=1/6$

This shows x and y are independent.


## 8

```{r}
tab <- matrix(c(1,2,3,4,5,6,7,3,1,3,2,3,3,2,1,1,3,1,2,2,1), ncol=3, byrow=FALSE)
colnames(tab) <- c('Particle ID','Energy Level','Time Period')
tab <- as.table(tab)
tab
```

### a

The probability distribution for the above data is given by the following table where X is columns and Y is the rows.

```{r}
tab <- matrix(c(1/7,2/7,1/7,0,0,2/7,0,0,1/7), ncol=3, byrow=TRUE)
colnames(tab) <- c('1','2','3')
rownames(tab) <- c('1','2','3')
tab <- as.table(tab)
addmargins(tab)
```

  
### b

The marginal distribution $p_1(x)$ is

```{r}
tab <- matrix(c(1/7,2/7,4/7), ncol=3, byrow=TRUE)
colnames(tab) <- c('1','2','3')
rownames(tab) <- c('p_1(x)')
tab <- as.table(tab)
tab
```


$p_1(x=1)=p_1(1)=1/7+0+0= 1/7 p_1(x=2)=p_1(2)=2/7+0+0= 2/7$ $p_1(x=3)p_1(3)=1/7+2/7+1/7= 4/7$

### c

The marginal distribution $p_2(y)$ is

```{r}
tab <- matrix(c(4/7,2/7,1/7), ncol=3, byrow=TRUE)
colnames(tab) <- c('1','2','3')
rownames(tab) <- c('p_2(y)')
tab <- as.table(tab)
tab
```


$p_2(y=1)=1/7+2/7+1/7 = 4/7$ $p_2(y=2)=0+0+2/7= 2/7$ $p_2(y=3)=0+0+1/7= 1/7$

### d

Find $p_2(y|x)$ for the data.

x=1
```{r}
tab <- matrix(c(1,0,0), ncol=3, byrow=TRUE)
colnames(tab) <- c('1','2','3')
rownames(tab) <- c('p_2(y|1)')
tab <- as.table(tab)
tab
```
x=2
```{r}
tab <- matrix(c(2/2,0,0), ncol=3, byrow=TRUE)
colnames(tab) <- c('1','2','3')
rownames(tab) <- c('p_2(y|2)')
tab <- as.table(tab)
tab
```
x=3
```{r}
tab <- matrix(c(1/4,2/4,1/4), ncol=3, byrow=TRUE)
colnames(tab) <- c('1','2','3')
rownames(tab) <- c('p_2(y|3)')
tab <- as.table(tab)
tab
```


$p(y|x=1)|(1/7)(4/7)=1/4| (2/7)/(4/7)=0.5 | (1/7)/(4/7)=0.25 p(y|x=2)| 0 | 0 |$ $(2/7)/(2/7)=1 p(y|x=3)| 0 | 0 | (1/7)/(1/7)=1$

## 9

### a

Integrate out and get $(e^{-y/10}/10$


### b

$E(Y)=\mu=\beta=10$

## 10

### a

If we take the double integral of $ce^{-x^2}$ with respect to y and then x, we get:

$c(y)e^{-x^2}$

then we use the bounds for y $(0\le y\le x)$ to get:

$c(x)e^{-x^2} - c(0)e^{-x^2} = c(x)e^{-x^2}$

then we use the bounds for x $(0\le x\le \infty)$ to get:

$-ce^{-x^2}/2$ from 0 to infinity.

Then we plug in the bounds and get (1/2)c=1, so c=2.

The whole z integral is equal to 1 on the interval by definition.

### b

$f_1(x)$ is equal to the integral of $f(x,y)$ with the 2 for c plugged in, thus: 

$f_1(x) = 2xe^{-x^2}$

We then take the integral of $f_1(x)$, using u-substitution, and get the whole thing is equal to 1.

### c

$ f_2(x,y) = f(x,y)/f_1(x)=(2e^{-x^2})/(2xe^{-x^2}) = 1/x$

## 11

$Cov(x,y)=E(XY)−E(X)E(Y)$

$E(XY)=∑_x∑_yxyp(x,y)=0$

$E(x)=∑_xxp_1(x)=0$

$E(y)=∑_yyp_2(y)=0$

Thus $Cov(X,Y)=0$, but $p(x,y)≠p_1(x)∗p_2(y)$. Therefore, X and Y are dependent.


## 12

### a

$(a,b)=(1,3)$

$E(Y)=(a+b)/2=2$

$E(Y) = 2$


```{r}
(1+3)/2
```

### b

$V(Y)=(a+b)/n=0.005555556$

```{r}
(1/3)/60
```


### c

```{r}
sqrt(0.005555556)

1-pnorm(2.5,2,0.0745356)

1-pnorm(1.5,2,0.0745356)

1-pnorm(1.5,2,0.0745356)-(1-pnorm(2.5,2,0.0745356))
```

The probability that $\bar y$ is between 1.5 ppm and 2.5 ppm is 1.

### d

```{r}
pnorm(2.2,2,0.0745356)
```

The probability that $\bar y$ exceeds 2.2 ppm is 0.9963548.

## 13

### a

```{r}
20*(.4)
```

$\mu = 8$

```{r}
sqrt(20*(.4)*(.6))
```

$\sigma = 2.19089$

```{r}
pnorm(1.5,8,2.19089)
```

The probability that fewer than 2 of 20 swordfish pieces exceeding the FDA limit is 0.001504432.


### b

```{r}
1-pnorm(10.5,8,2.19089)
```

The probability that more then half of the 20 pieces of swordfish have mercury levels exceeding the FDA limit is 0.1269165.

### c

```{r}
pbinom(1,20,.4)

1-pbinom(10,20,.4)
```

From the values we found in parts a and b the normal normal approximation provides a good estimation of the binomial.

## 14

```{r}
leadcop<-read.csv("LEADCOPP.csv")
```

### a

```{r}
t.test(leadcop$LEAD,conf.level = 0.99)
```

The mean lead level 99 percent confidence interval is (-1.15,6.92).

### b

```{r}
t.test(leadcop$COPPER,conf.level = 0.99)
```

The mean copper level 99 percent confidence interval is (0.1519,0.6647).

### c

Essentially we are 99 percent confident that the mean values for lead and copper levels are found inside the intervals that we found.

### d

99 percent confidence means that in repeated sampling 99 percent of the intervals constructed in the same way will contain the actual mean.

## 15

```{r}
sol<-read.csv("SOLARAD.csv")
names(sol)
with(sol,t.test(STJOS,IOWA,paired=TRUE,conf.level=0.95))
```

We can be 95 percent confident that the true mean difference of the two locations lies in the interval (156.84,239.18).

## 16

```{r}
dia<-read.csv("DIAZINON.csv")
with(dia,t.test(DAY,NIGHT,paired=TRUE,conf.level=0.90))
```

### a 

We are 90 percent certain the true mean difference lies between -58.894 and -18.92.

### b

We must assume the populations are distributed normally.

### c

Because the confidence interval does not contain zero there is evidence that the mean differs greatly in the day and night compared to one another.
